Whose Truth is it Anyway? An Experiment on Annotation Bias

### 2023-06-22 report/draft.Rmd
update references

Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020penguin].
Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
Datasets are always "a worldview" [26] and, as such, data always remains biased." (Miceli et al., 2022, p. 5). This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators. The answer of the annotators in turn influence how machine learning examples classify new models. For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance[@gordon2022, @Larimore, @waseem2016]. Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded, and that notably that African American English dialects are often misunderstood by such models.

>>>>

Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020penguin].
Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
Datasets are always "a worldview" [26] and, as such, data always remains biased." (Miceli et al., 2022, p. 5). This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators. The answer of the annotators in turn influence how machine learning examples classify new models. For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance[@gordon2022, @larimore2021reconsidering, @waseem2016]. Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded, and that notably that African American English dialects are often misunderstood by such models.
